---
title: "hw4_problem1"
author: "Feben Asefaha"
date: "10/25/2017"
output: html_document
---
Feben Asefaha (fta2112) is the lead for Problem 1, and Nadiya Pavlishyn (np2547) is the collaborator.

###Problem 1
```{r load libraries and data}

library(tidyverse)
library(readr)
library(readxl)
library(haven)
library(janitor)

#loading data, tidying
pups_data <- read_csv("./FAS_pups.csv") %>% 
  clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))

#gathering variables and omitting blank observations

pups_data_tidy = gather(pups_data, key = pd_measure, value = pd_outcome, pd_ears:pd_walk) %>% filter(!is.na(pd_outcome))

knitr::opts_chunk$set(echo = TRUE)
```



```{r}
#creating plot
library(ggridges)

#ridge plot to look across day differences in developemnt
pups_plot = pups_data_tidy%>%
  ggplot(aes(x = pd_outcome, y = pd_measure)) + geom_density_ridges(scale = .85, aes(fill = pd_measure), alpha = .4) + labs(title = "Pup Development Across Days",
       x = "Number of Days",
       y = "Development Point")

pups_plot
```


The pups_data dataset contained `r length(pups_data$litter_number)` observations across `r length(pups_data)` variables originally.  It was then tidied for our purposes in this problem.  The developmental points: ears, eyes, pivot, and walk were spread out for us to be able to evaluate the difference in time among the different points.  The resulting dataset had `r length(pups_data_tidy$litter_number)` obersvations across `r length(pups_data_tidy)` variables.  We made a ridge plot showing the differences in time among these developmental points. For most points, the number of days seems to vary, but it contained to a certain time frame.  It seeems for pd_eyes, the majority of pups develop eyes in the day 12-14 time frame, with very little developing before or after.  For the others, this time interval seems to be a little more spread out acorss time, and spread evenly through the litters. For example, a pretty even amount of pups seem to develop at pd_walk in the day 8-12 time frame.